A Multi-telescope Campaign on \frb: Implications for the FRB Population

### 2017-05-19 repeater-vladata.tex
through cees comments

\author{C.~G.~Bassa}
\affiliation{ASTRON, Netherlands Institute for Radio Astronomy, Postbus 2, 7990 AA, Dwingeloo, The Netherlands}

@@

\author{J.~W.~T.~Hessels}
\affiliation{ASTRON, Netherlands Institute for Radio Astronomy, Postbus 2, 7990 AA, Dwingeloo, The Netherlands}
\affiliation{Anton Pannekoek Institute for Astronomy, University of Amsterdam, Science Park 904, 1098 XH Amsterdam, The Netherlands}

@@

\author{D.~Michilli}
\affiliation{ASTRON, Netherlands Institute for Radio Astronomy, Postbus 2, 7990 AA, Dwingeloo, The Netherlands}
\affiliation{Anton Pannekoek Institute for Astronomy, University of Amsterdam, Science Park 904, 1098 XH Amsterdam, The Netherlands}

@@

\begin{abstract}
We present results of the coordinated observing campaign that made the first subarcsecond localization of a Fast Radio Burst, \frb. During this campaign, we made the first detection of an FRB burst by multiple telescopes: the VLA at 3~GHz and the Arecibo Observatory at 1.4~GHz. Of the nine bursts detected by the Very Large Array at 3~GHz, four had simultaneous observing coverage at other observatories. 
We use multi-observatory constraints and modeling of bursts seen only at 3~GHz to confirm earlier results showing that burst spectra are not well characterized by a power law. We find that burst spectra are characterized by a $\sim500$~MHz envelope and apparent radio energy as high as $10^{40}$\ erg. We measure significant changes in the apparent dispersion between bursts that can be attributed to frequency-dependent profiles or some other intrinsic burst structure that adds a random error to the estimate of DM by up to 1\%. We use \frb\ as a prototype of the FRB class to estimate a volumetric birth rate of FRB sources $R_{\rm{FRB}} \approx 5\times10^{-5}/N_r$\ Mpc$^{-3}$\ yr$^{-1}$, where $N_r$\ is the number of bursts per source over its lifetime. This rate is broadly consistent with models of FRBs from young pulsars or magnetars born in superluminous supernovae or long gamma-ray bursts, if the typical FRB repeats on the order of $10^3$\ times during its lifetime.

\end{abstract}

@@

Fast Radio Bursts (FRBs) were discovered ten years ago with the detection of a millisecond-duration radio transient with an anomalously high dispersion measure \citep{2007Sci...318..777L}. The largest measured DMs imply that FRBs originate outside of our Galaxy, potentially at cosmological distances, and are orders of magnitude more luminous than pulses from Galactic pulsars \citep{2013Sci...341...53T}. There are now 21 FRBs publicly known \footnote{See \url{http://www.astronomy.swin.edu.au/pulsar/frbcat}.} with dispersion measures (DMs) as high as 1600 pc cm$^{-3}$ and intrinsic temporal widths of order milliseconds. Both their energetics and distance have inspired a wide variety of models and astrophysical applications \citep[e.g.,][]{2014ApJ...780L..33M, 2014ApJ...797...70K, 2015MNRAS.450L..71F, 2016MNRAS.458L..19C, 2016MNRAS.457..232C, 2016MNRAS.462L..16P}. However, that potential was limited by the lack of a definitive association of an FRB to an extragalactic host.

@@

Beginning in August of 2015 (MJD 57623), we made the first of nine detections of \frb\ with the Karl G.\ Jansky Very Large Array \citep[VLA;][]{LOC} and localized it with a precision of 0.1\arcsec. Deep optical observing with the Gemini Observatory associated the FRB with a host galaxy at $z=0.193$\ \citep{OPT} and the European VLBI Network detected four more bursts to localize the source with a precision of 0.01\arcsec, four orders of magnitude better than any other FRB \citep[precision of $\sim40$\ pc in linear distance;][]{EVN}. The lookback and luminosity distances for \frb\ are 746 and 972 Mpc, respectively, assuming a concordance cosmology with parameters given by \citet{2016A&A...594A..13P}. If \frb\ is representative of all FRBs, then we should expect them to be useful as probes of the intergalactic medium and their host galaxies. Thus, the confirmation of a cosmological distance for \frb\ has begun to fulfil the promise implied by the first FRB detection.

Many new models for the origin of FRBs have been developed in response to the localization of \frb\ \citep{2017arXiv170104815K, 2017arXiv170102370M, 2017arXiv170104094Z, 2017arXiv170102492D, 2017arXiv170208644B, 2017arXiv170300393T}. The association of \frb\ with a compact, persistent radio source is consistent with bursts coming from a young magnetar that powers a luminous pulsar wind nebula \citep{2017arXiv170104815K}. At the same time, \frb\ is found in a low-metallicity dwarf galaxy \citep{OPT, 2017arXiv170102370M}, which was not predicted by origin models that scale with stellar mass or star formation \citep{2017arXiv170400022N}. These galaxies are the preferred environment for long GRBs and Type-I superluminous supernovae (LGRB and SLSN-I, respectively), which have been suggested are signatures of the magnetar birth \citep{2008AJ....135.1136M, 2014ApJ...787..138L}.

@@

Table \ref{tab:spec} summarizes the spectrotemporal properties of all nine VLA bursts. Parameters such as integrated flux density, S/N, and Stokes V were measured from the burst properties integrated in frequency. Stokes V values were estimated from the dual-circular feeds as $(RR-LL)/(RR+LL)\approx3\%$, which is comparble to the systematic error expected from beam squint at this location in the primary beam \footnote{See EVLA memo \#195 at \url{https://library.nrao.edu/public/memos/evla/EVLAM_195.pdf}.}. Given that systematic effects dominate the apparent circular polarization, this observation limits the fractional circular polarization to less than 3\%.

@@

The dynamic radio spectra (time versus frequency) were modeled using a Markov chain Monte Carlo (MCMC) method. We used the Goodman and Weare affine invariant sampler \citep{goodman2010ensemble} as implemented in \emph{emcee} \citep{2013PASP..125..306F}. The dynamic spectra were modeled prior to dispersion correction. The spectral structure was modeled as a broad Gaussian envelope as a function of channel ``ch'' as:
\begin{equation}
G(\rm{ch}) = \frac{\rm{amp}}{\sqrt{2\pi}} \exp{\left[-\frac{1}{2} \left(\frac{\rm{ch} - \rm{center}}{\sigma_{ch}}\right)^2\right]}
\end{equation}
\noindent where ``amp'' is in mJy per 5~ms integration, ``center'' is the Gaussian center channel, and $\sigma_{ch}$\ is its width. The arrival time was modeled with a cold plasma dispersion law with the arrival time in units of integrations of:
\begin{equation}
\rm{i} (\rm{ch}) = \rm{i0} + 4.1488\times10^{-3}\ \rm{DM} \left[f(\rm{ch})^{-2} - f(\rm{ch0})^{-2}\right]/(5 \rm{ms})
\end{equation}
\noindent where ``i0'' is the arrival time at the highest frequency, ``ch0'', and function ``f'' maps the channel to a frequency in units of GHz. Finally, the pulses were assumed to have an intrinsically square temporal width (parameter $W_{int}$). This 6-parameter model defined the intrinsic signal that was then distributed over a fixed time-frequency observing grid.

@@

The integrated burst flux density is calculated by integrating the burst spectral model (Table \ref{tab:spec}) in frequency and time. Some VLA burst spectra seem to be contained by the 2.5 to 3.5~GHz band and most of them seem to have Gaussian envelopes that are well modeled by the emission within that band. Assuming that the Gaussian shape defines the full emission window, the integrated 3~GHz flux density can be converted to a total energy with no further assumptions about the burst spectral properties as:
\begin{equation}
E_{\rm{int}} = (\rm{amp}\ 5\ \rm{ms}) ( \sigma_{ch}\ 2.355 4\ \rm{MHz}) 10^{-23} (4\pi L_d^2)\ \rm{erg}
\end{equation}
\noindent where the first term represents the burst fluence, the second term represents the integral of the burst spectrum, and $L_d$\ is the luminosity distance to the source.

@@

After correcting for barycentric and typical dispersion delays, the burst spectra at the VLA and at Arecibo show a residual temporal drift. This drift can be interpreted as an excess DM, but contemporaneous observations at 1.4~GHz measured a significantly lower DM (560.5 pc cm$^{-3}$; Hessels et al, in prep). This excess DM is comparable to the burst-to-burst variation in DM within the whole sample of VLA 3~GHz bursts. This variation had been noted in earlier 1.4~GHz Arecibo observations of \frb\ \citep{2016arXiv160308880S}, but the VLA bursts demonstrate that this burst-specific DM is seen simltaneously from 1.4 to 3~GHz.

@@

\section*{Acknowledgements}
% observatories
The National Radio Astronomy Observatory is a facility of the National Science Foundation operated under cooperative agreement by Associated Universities, Inc..
Part of this research was carried out at the Jet Propulsion Laboratory, California Institute of Technology, under a contract with the National Aeronautics and Space Administration.
% systems
This research made use of Astropy, a community-developed core Python package for Astronomy (Astropy Collaboration, 2013).
%individuals
CJL is supported by the University of California Office of the President under Lab Fees Research Program Award 237863 and NSF award 1611606. MAM is supported by NSF award 1458952. 
KPM's research is supported by the Oxford Centre for Astrophysical Surveys which is funded through the Hintze Family Charitable Foundation. AS gratefully acknowledges support from the European Research Council under grant ERC-2012- StG-307215 LODESTONE. The AMI-LA telescope gratefully acknowledges support from the European Research Council under grant ERC-2012- StG-307215 LODESTONE, the UK Science and Technology Facilities Council (STFC) and the University of Cambridge. 
PS holds a Covington Fellowship at DRAO. 
The LWA1 station is supported by the National Science Foundation under grant 1139974 of the University Radio Observatory program.
VMK acknowledges support from NSERC, CIFAR, the Canada Research Chair Program and from the Lorne Trottier Chair in Astrophysics \& Cosmology.

We thank the VLA staff for their support of \rf\ development and Liam Connor for useful feedback. We acknowledge partial support from the Research Corporation for Scientific Advancement (RCSA) for participation in the meeting Fast Radio Bursts: New Probes of Fundamental Physics and Cosmology at the Aspen Center for Physics (February 12-17, 2017).

>>>>

\author{C.~G.~Bassa}
\affiliation{ASTRON, The Netherlands Institute for Radio Astronomy, Postbus 2, NL-7990 AA, Dwingeloo, The Netherlands}

@@

\author{J.~W.~T.~Hessels}
\affiliation{ASTRON, The Netherlands Institute for Radio Astronomy, Postbus 2, NL-7990 AA, Dwingeloo, The Netherlands}
\affiliation{Anton Pannekoek Institute for Astronomy, University of Amsterdam, Science Park 904, 1098 XH Amsterdam, The Netherlands}

@@

\author{D.~Michilli}
\affiliation{ASTRON, The Netherlands Institute for Radio Astronomy, Postbus 2, NL-7990 AA, Dwingeloo, The Netherlands}
\affiliation{Anton Pannekoek Institute for Astronomy, University of Amsterdam, Science Park 904, 1098 XH Amsterdam, The Netherlands}

@@

\begin{abstract}
We present results of the coordinated observing campaign that made the first subarcsecond localization of a Fast Radio Burst, \frb. During this campaign, we made the first detection of an FRB burst by multiple telescopes: the VLA at 3~GHz and the Arecibo Observatory at 1.4~GHz. Of the nine bursts detected by the Very Large Array at 3~GHz, four had simultaneous observing coverage at other observatories. We use multi-observatory constraints and modeling of bursts seen only at 3~GHz to confirm earlier results showing that burst spectra are not well modeled by a power law. We find that burst spectra are characterized by a $\sim500$~MHz envelope and apparent radio energy as high as $10^{40}$\ erg. We measure significant changes in the apparent dispersion between bursts that can be attributed to frequency-dependent profiles or some other intrinsic burst structure that adds a random error to the estimate of DM by up to 1\%. We use \frb\ as a prototype of the FRB class to estimate a volumetric birth rate of FRB sources $R_{\rm{FRB}} \approx 5\times10^{-5}/N_r$\ Mpc$^{-3}$\ yr$^{-1}$, where $N_r$\ is the number of bursts per source over its lifetime. This rate is broadly consistent with models of FRBs from young pulsars or magnetars born in superluminous supernovae or long gamma-ray bursts, if the typical FRB repeats on the order of $10^3$\ times during its lifetime.
\end{abstract}

\keywords{stars: neutron, radio continuum: stars, techniques: interferometric, supernovae: general}

@@

Fast Radio Bursts (FRBs) were discovered ten years ago with the detection of a millisecond-duration radio transient with an anomalously high dispersion measure \citep{2007Sci...318..777L}. The largest measured DMs imply that FRBs originate outside of our Galaxy, potentially at cosmological distances, and are orders of magnitude more luminous than pulses from Galactic pulsars \citep{2013Sci...341...53T}. There are now 22 FRBs publicly known \footnote{See \url{http://www.astronomy.swin.edu.au/pulsar/frbcat} \citep{2016PASA...33...45P}.} with dispersion measures (DMs) as high as 1600 pc cm$^{-3}$ and intrinsic temporal widths of order milliseconds. Both their energetics and distance have inspired a wide variety of models and astrophysical applications \citep[e.g.,][]{2014ApJ...780L..33M, 2014ApJ...797...70K, 2015MNRAS.450L..71F, 2016MNRAS.458L..19C, 2016MNRAS.457..232C, 2016MNRAS.462L..16P}. However, that potential was limited by the lack of a definitive association of an FRB to an extragalactic host.

@@

Beginning in August 2016 (MJD 57623), we made the first of nine detections of \frb\ with the Karl G.\ Jansky Very Large Array \citep[VLA;][]{LOC} and localized it with a precision of 0\farcs1. Deep optical observations with the Gemini Observatory associated the FRB with a host galaxy at $z=0.193$\ \citep{OPT} and the European VLBI Network detected four more bursts to localize the source with a precision of 0.01\arcsec, four orders of magnitude better than any other FRB \citep[precision of $\sim40$\ pc in linear distance;][]{EVN}. The lookback and luminosity distances for \frb\ are 746 and 972 Mpc, respectively, assuming a concordance cosmology with parameters given by \citet{2016A&A...594A..13P}. If \frb\ is representative of all FRBs, then we should expect them to be useful as probes of the intergalactic medium and their host galaxies. Thus, the confirmation of a cosmological distance for \frb\ has begun to fulfil the promise implied by the first FRB detection.

Many new models for the origin of FRBs have been developed in response to the localization of \frb\ \citep{2017arXiv170104815K, 2017arXiv170102370M, 2017arXiv170104094Z, 2017arXiv170102492D, 2017arXiv170208644B, 2017arXiv170300393T}. The association of \frb\ with a compact, persistent radio source is consistent with bursts coming from a young magnetar that powers a luminous pulsar wind nebula \citep{2017arXiv170104815K}. At the same time, \frb\ is found in a low-metallicity dwarf galaxy \citep{OPT}, which was not predicted by origin models that scale with stellar mass or star formation \citep{2017arXiv170400022N}. These galaxies are the preferred environment for long GRBs and Type-I superluminous supernovae (LGRB and SLSN-I, respectively), which have been suggested are signatures of the magnetar birth \citep{2008AJ....135.1136M, 2014ApJ...787..138L}.

@@

Table \ref{tab:spec} summarizes the spectrotemporal properties of all nine VLA bursts. Parameters such as integrated flux density, S/N, and Stokes V were measured from the burst properties integrated in frequency. Stokes V values were estimated from the dual-circular feeds as $(RR-LL)/(RR+LL)\approx3\%$, which is comparable to the systematic error expected from beam squint at this location in the primary beam \footnote{See EVLA memo \#195 at \url{https://library.nrao.edu/public/memos/evla/EVLAM_195.pdf}.}. Given that systematic effects dominate the apparent circular polarization, this observation limits the fractional circular polarization to less than 3\%.

@@

The dynamic radio spectra (time versus frequency) were modeled using a Markov chain Monte Carlo (MCMC) method. We used the Goodman and Weare affine invariant sampler \citep{goodman2010ensemble} as implemented in \emph{emcee} \citep{2013PASP..125..306F}. The dynamic spectra were modeled prior to dispersion correction. The spectral structure was modeled as a broad Gaussian envelope as:
\begin{equation}
G(\nu) = \frac{\rm{A}}{\sqrt{2\pi} \sigma} \exp{\left[-\frac{1}{2} \left(\frac{\nu - \nu_0}{\sigma}\right)^2\right]}
\end{equation}
\noindent where ``A'' is in mJy per 5~ms integration, $\nu_0$\ is the Gaussian center frequency, and $\sigma$\ is its width. The arrival time was modeled with a cold plasma dispersion law with the arrival time in units of integrations of:
\begin{equation}
\rm{i} (\nu) = \rm{i0} + 4.1488\times10^{-3}\ \rm{DM} \left(\nu^{-2} - \nu_0^{-2}\right)/(5 \rm{ms})
\end{equation}
\noindent where ``i0'' is the arrival integration at the highest frequency and $\nu$\ is in units of GHz. Finally, the pulses were assumed to have an intrinsically square temporal width (parameter $W_{int}$). These 6 parameters defined the intrinsic signal that was then distributed over a fixed time-frequency observing grid.

@@

The integrated burst flux density is calculated by integrating the burst spectral model (Table \ref{tab:spec}) in frequency and time. Some VLA burst spectra seem to be contained by the 2.5 to 3.5~GHz band and most of them seem to have Gaussian envelopes that are well modeled by the emission within that band. Assuming that the Gaussian shape defines the full emission window, the integrated 3~GHz flux density can be converted to a total energy with no further assumptions about the burst spectral properties as:
\begin{equation}
E_{\rm{int}} = (\rm{A}\ 5\ \rm{ms}) ( \sigma_{ch}\ 2.355 4\ \rm{MHz}) 10^{-23} (4\pi L_d^2)\ \rm{erg}
\end{equation}
\noindent where the first term represents the burst fluence, the second term represents the integral of the burst spectrum, 2.355 scales $\sigma$\ to FWHM, and $L_d$\ is the luminosity distance to the source.

@@

After correcting for barycentric and typical dispersion delays, the VLA+AO burst spectrum has a residual temporal drift. This drift can be interpreted as an excess DM, but contemporaneous observations at 1.4~GHz measured a significantly lower DM (560.5 pc cm$^{-3}$; Hessels et al, in prep). This excess DM is comparable to the burst-to-burst variation in DM within the whole sample of VLA 3~GHz bursts. This variation had been noted in earlier 1.4~GHz Arecibo observations of \frb\ \citep{2016arXiv160308880S}, but the VLA bursts demonstrate that this burst-specific DM is seen simltaneously from 1.4 to 3~GHz.

@@

\section*{Acknowledgements}
We thank the VLA staff for their support of \rf\ development and Liam Connor for useful feedback. We acknowledge partial support from the Research Corporation for Scientific Advancement (RCSA) for participation in the meeting Fast Radio Bursts: New Probes of Fundamental Physics and Cosmology at the Aspen Center for Physics (February 12-17, 2017).

% observatories
The National Radio Astronomy Observatory is a facility of the National Science Foundation operated under cooperative agreement by Associated Universities, Inc..
Part of this research was carried out at the Jet Propulsion Laboratory, California Institute of Technology, under a contract with the National Aeronautics and Space Administration.
% systems
This research made use of Astropy, a community-developed core Python package for Astronomy (Astropy Collaboration, 2013).
%individuals
CJL is supported by the University of California Office of the President under Lab Fees Research Program Award 237863 and NSF award 1611606. MAM is supported by NSF award 1458952. 
JWTH and CGB acknowledge support from the European Research Council under the European Union's Seventh Framework Programme (FP/2007-2013) / ERC Grant Agreement nr.\ 337062 (DRAGNET; PI Hessels).
KPM's research is supported by the Oxford Centre for Astrophysical Surveys which is funded through the Hintze Family Charitable Foundation. AS gratefully acknowledges support from the European Research Council under grant ERC-2012- StG-307215 LODESTONE. The AMI-LA telescope gratefully acknowledges support from the European Research Council under grant ERC-2012- StG-307215 LODESTONE, the UK Science and Technology Facilities Council (STFC) and the University of Cambridge. 
PS holds a Covington Fellowship at DRAO. 
The LWA1 station is supported by the National Science Foundation under grant 1139974 of the University Radio Observatory program.
VMK acknowledges support from NSERC, CIFAR, the Canada Research Chair Program and from the Lorne Trottier Chair in Astrophysics \& Cosmology.

\facility{EVLA, Arecibo, Effelsberg, LWA, AMI}
